Network based multifactorial modelling microRNA-target interaction.

### 2019-09-10 article.tex
news to article

  \title{Network based multifactorial modelling microRNA-target interaction.}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Selcen Ari\footnote{Department of Bioengineering, Yildiz Technical
  University, Istanbul, Turkey} \\ Alper Yilmaz\footnote{Department of Bioengineering, Yildiz Technical
  University, Istanbul, Turkey}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{21 Ağustos 2019}

@@

MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key
regulator of gene expression through various post-transcriptional
mechanisms. Although the mechanisms by which miRNA represses are not
fully understood, miRNAs predominantly repress their targets. Repressive
activities of miRNAs vary depending on many factors that are significant
to microRNA:target interactions. These factors include microRNA:target
binding energy, binding location in target sequence, base pairing types
between miRNA and target, abundance of miRNAs and targets (Grimson et
al. \protect\hyperlink{ref-grimson_microrna_2007}{2007}). Binding
energies of miRNA:target complexes vary based on nucleotide context and
determine folding stability of complex (Cao and Chen
\protect\hyperlink{ref-cao_predicting_2012}{2012}). It has been
demonstrated that the binding energy between miRNA and target indicates
stability or affinity of complex (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}) and does not directly
determine repressive activity of miRNA (Cao and Chen
\protect\hyperlink{ref-cao_predicting_2012}{2012}). Early studies have
argued that 2-8 nt sequence, seed, located in miRNA 5'end bind to
specific sequence located in 3'UTR of its target (Bartel
\protect\hyperlink{ref-bartel_micrornas_2004}{2004}; Lewis, Burge, and
Bartel \protect\hyperlink{ref-lewis_conserved_2005}{2005}). In recent
studies, it has been shown that miRNAs can interact with targets via
sequences located in regions such as 5'UTR or CDS (Hausser et al.
\protect\hyperlink{ref-hausser_analysis_2013}{2013}; Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). These studies
also showed that binding location could indicate functionality of
miRNA:target interaction or be effective on abundance of targets(TOFIX
abundance etkisi derken, bazen azaltabilir, bazen arttırabilir,
manasında mı? answer: evet hocam anlaşılmıyorsa değiştirebilirim). It
has been shown that miRNAs exhibit repressive activity via, 6-8 nt long
sequence that is perfectly complementary with targets, seed region at
the 5' end of miRNAs (Bartel
\protect\hyperlink{ref-bartel_micrornas:_2009}{2009}; Grimson et al.
\protect\hyperlink{ref-grimson_microrna_2007}{2007}). On the other hand,
some researchers have reported that seed sequence of miRNA can have
mismatches or bulged/wobble nucleotides and may locate in region other
than 5'end of miRNAs (Hafner et al.
\protect\hyperlink{ref-hafner_transcriptome-wide_2010}{2010}; Helwak et
al. \protect\hyperlink{ref-helwak_mapping_2013}{2013}). On top of all
these factors, abundance of miRNAs and targets and miRNA:target ratio in
cells predominantly affect efficiency of miRNA:target interaction (Arvey
et al. \protect\hyperlink{ref-arvey_target_2010}{2010}; Bosson, Zamudio,
and Sharp \protect\hyperlink{ref-bosson_endogenous_2014}{2014}; Denzler
et al. \protect\hyperlink{ref-denzler_assessing_2014}{2014}).

As it is possible for miRNAs to suppress multiple targets, an individual
mRNA molecule can also be targeted by multiple miRNAs. In that case, the
targeted mRNAs exhibit competitor behavior, that is hypothesized as
competing endogenous RNAs (ceRNAs) (Ala et al.
\protect\hyperlink{ref-ala_integrated_2013}{2013}; Cesana and Daley
\protect\hyperlink{ref-cesana_deciphering_2013}{2013}), against their
miRNAs. Briefly, Ala et al.~have explained the ceRNA hypothesis as
disturbance of the other target when one of the targets on a
steady-state system that included one miRNA and two target was perturbed
with expression change (Ala et al.
\protect\hyperlink{ref-ala_integrated_2013}{2013}). Regarding
interaction between miRNAs and their target in a cell, explaining and
predicting results of an individual perturbation is difficult due to
complexity of interactions. Various computational and experimental
studies have tackled the problem of unraveling ceRNA:miRNA interactions.
It has been observed that when abundance of one of the targets of
miR-122 was increased, the other target expression also slightly
increased as a result of decreasing repressive activity of miR-122 on
them (Denzler et al.
\protect\hyperlink{ref-denzler_assessing_2014}{2014}). Bosson et
al.~have developed a mathematical model for changes on total target pool
concentration after grouping targets according to affinity and
demonstrated that miRNA activity correlated with affinity between miRNA
and target (Bosson, Zamudio, and Sharp
\protect\hyperlink{ref-bosson_endogenous_2014}{2014}). Cooperative
efficiency of miRNAs as well as competitor behaviors of targets were
also studied and it has been demonstrated to be crucial for regulating
available mRNA levels of targets (Denzler et al.
\protect\hyperlink{ref-denzler_impact_2016}{2016}). MiRNA:target
interactions have been modeled as stoichiometric and catalytic
mechanisms and Figliuzzi et al.~have recommended handling models in
network context (Figliuzzi, Marinari, and De Martino
\protect\hyperlink{ref-figliuzzi_micrornas_2013}{2013}). The model that
can explain miRNA target interaction through topological features has
been applied at bipartite network by Nitzan et al.~(Nitzan et al.
\protect\hyperlink{ref-nitzan_interactions_2014}{2014}). Robinson and
Henderson applied the model that handles miRNA:target direct and
indirect interactions via common miRNA of gene and target of miRNA, at
bipartite network. It has been demonstrated that all miRNAs and targets
in the network can interact with each other through common miRNAs and
genes, without interaction between the same type of nodes (Robinson and
Henderson
\protect\hyperlink{ref-robinson_modelling_2018}{2018}).Associated genes
that are targets of the same miRNAs have been found with help of
correlation of gene expression changes in recent algorithm (Markus List
\protect\hyperlink{ref-markus_list_sponge_2017}{2017}). List et al.~have
specified that their approach can be useful for ceRNA studies and
published their approach as an R package.

@@

Initially, the network is assumed in steady-state (\autoref{fig1}a)
condition and needs a trigger for initiating calculations. The trigger
can be a change in expression level of one or more genes
(\autoref{fig1}b). After a trigger, the network undergoes iterative
cycle of calculations at each of which distribution of miRNA in local
neighborhood is recalculated (\autoref{fig1}c). Based on new miRNA
distribution, expression level of each node (i.e.~ceRNA) is updated. Due
to nodes which are targeted by multiple miRNAs, the change in one
neighborhood spreads to other neighborhoods (\autoref{fig1}d),
consequently have potential to effect whole network due to ``ripple
effect''.

@@

We have applied our model in a real dataset for which experimental
measurements of various factors were available. Expression levels of
miRNA and genes in tumor and normal tissue of single patient are
retrieved from
\href{https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga}{TCGA}(\#TODO).
High-throughput experimental datasets which are provided miRNA:gene
target pairs with interaction factors (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). We have combined
miRNA and gene expression datasets via miRNA:target gene dataset
retrieved from \ldots{} (\#FIX name of database and its citation)
answer-\textgreater{}suplementary datasets of (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). Detailed
description of network construction and its code is available in
Supplementary data (\#TODO burada link nasıl olacak, word dosyası veya
PDF dosyası ismi mi vermemiz mi gerekiyor, yoksa sadece Sup Data
denilmesi yeterli mi? ilgili dosya: TCGA\_E9-A1N5\_article.Rmd). ABCC1
gene, over-expressed in tumor tissue, was selected as trigger for
simulation of integrated dataset. After simulation of network, we have
compared simulation results and tumor tissue expression levels.

@@

We have developed a network-based approach to assess effects of
expression level changes in competitive ceRNA regulation. The basic mode
of miRNA repression activity has been based on miRNA and target
abundance in various researches(Arvey et al.
\protect\hyperlink{ref-arvey_target_2010}{2010}; Denzler et al.
\protect\hyperlink{ref-denzler_assessing_2014}{2014}). Our approach can
effortlessly calculate effects of expression changes when abundance
levels of miRNAs and targets is only available factor. In sample network
given in Figure (\autoref{fig1}), after an increase in expression level
of a gene (G2), expression values of other genes also changed due to
redistribution of miRNA among its targets. Previous studies have shown
that if a gene abundance increases in ceRNA system, expression levels of
genes targeted by shared miRNA are also affected (Lai, Wolkenhauer, and
Vera \protect\hyperlink{ref-lai_understanding_2016}{2016}; Salmena et
al. \protect\hyperlink{ref-salmena_cerna_2011}{2011}; Tay, Rinn, and
Pandolfi \protect\hyperlink{ref-tay_multilayered_2014}{2014}). It was
observed that expression levels of primary neighborhoods which do not
interact with another miRNA of the trigger gene change in relation to
their expression. However, expression level change of common target (G4)
in network has been observed different from G1 that has the same
expression value with G4 target at initial conditions because more than
one miRNA repress this target (\autoref{fig1}d) (\#FIX last sentence
needs some clarification, after ``but cooperative\ldots{}'').

@@

In regulation of miRNA target sample system (\autoref{fig3}), miRNA (M1)
repressive efficiency on the primary triggering gene (G2) is low in
steady-state. So it has been observed that the regulatory activities of
miRNA efficiencies on the targets are weak after the increase of miRNA
target G2. When model is triggered with two fold increase in expression
level of common target (G4), the changes of other gene expressions have
observed more prominent \href{}{}. Furthermore, change in expression
level of target gene that has strong miRNA repression efficiency
resulted in evident perturbation in network. On the other hand, it was
observed that Gene2 was weakly affected from change of Gene4 expression
because of its weak interaction factors. Since perturbation efficiency
of each gene is different, we have developed a function which screens
each gene in the network for their perturbation efficiencies.
\textbf{(see details in Supplementary File(fig1\_2app.Rmd)}. On the
other hand, when we applied the method on the minimal dataset, Gene4 has
been found to be the most efficient element in terms of number of
perturbed elements and miRNA2 (M2) has been found to be causing the
highest mean in expression changes. {[}{]}{[}\#TODO2{]}

\href{}{} We used our model to simulate a real dataset which contains
thousands of genes and hundreds of miRNAs. Our model can successfully
simulate perturbations in such large network despite complex behaviors
and struggle to reach steady-state \textbf{(see details in Supplementary
File(TCGA\_E9-A1N5\_article.Rmd)}. Simulations show that change in
expression level of single gene has potential to effect whole network,
perturbing almost all nodes. These observations are in accordance with
competing endogenous RNA hypothesis where genes targeted with many
common miRNAs subsequently transmit perturbation to neighboring groups.

Network based approaches for analyzing miRNA:target interactions have
been developed in earlier studies. An initial attempt demonstrated the
ceRNA crosstalk in a network-like minimal interaction structure with
concentrations of ceRNA and miRNAs (Figliuzzi, Marinari, and De Martino
\protect\hyperlink{ref-figliuzzi_micrornas_2013}{2013}). Next, a network
based kinetic model integrating miRNA and target rates of transcription,
degradation, binding and unbinding was developed (Nitzan et al.
\protect\hyperlink{ref-nitzan_interactions_2014}{2014}) using high
throughput experimental dataset about miR-92a depletion (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}). In this study, it
was demonstrated that distant ceRNAs can interact with each other via
indirect links, and the interactions are effected depending on distance
between ceRNAs or topological features of network (Nitzan et al.
\protect\hyperlink{ref-nitzan_interactions_2014}{2014}). More recently,
an approach to detect ceRNA interaction by using the miRNA expression,
gene expression and common miRNAs between gene targets was developed
(Markus List \protect\hyperlink{ref-markus_list_sponge_2017}{2017})
which was effective in analyzing genes through miRNAs. Based on the
observations that a miRNA can exhibit strong functionality to a target
but may not against an other, the authors have concluded that common
miRNA based approach may not be suitable for understanding regulations
of ceRNA interactions.

@@

@@

@@

@@

@@

>>>>

  \title{Network based multifactorial modelling microRNA-target interaction.}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Selcen Ari\footnote{Department of Bioengineering, Yildiz Technical
  University, Istanbul, Turkey} \\ Alper Yilmaz\footnote{Department of Bioengineering, Yildiz Technical
  University, Istanbul, Turkey}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{10 Eylül 2019}

@@

MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key
regulator of gene expression through various post-transcriptional
mechanisms. Although the mechanisms by which miRNA represses are not
fully understood, miRNAs predominantly repress their targets. Repressive
activities of miRNAs vary depending on many factors that are significant
to miRNA:target interactions. These factors include miRNA:target binding
energy, binding location in target sequence, base pairing types between
miRNA and target, abundance of miRNAs and targets (Grimson et al.
\protect\hyperlink{ref-grimson_microrna_2007}{2007}). Binding energies
of miRNA:target complexes vary based on nucleotide context and determine
folding stability of complex (Cao and Chen
\protect\hyperlink{ref-cao_predicting_2012}{2012}). It has been
demonstrated that the binding energy between miRNA and target indicates
stability or affinity of complex (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}) and does not directly
determine repressive activity of miRNA (Cao and Chen
\protect\hyperlink{ref-cao_predicting_2012}{2012}). Early studies have
argued that 2-8 nt sequence, seed, located in miRNA 5'end bind to
specific sequence located in 3'UTR of its target (Bartel
\protect\hyperlink{ref-bartel_micrornas_2004}{2004}; Lewis, Burge, and
Bartel \protect\hyperlink{ref-lewis_conserved_2005}{2005}). In recent
studies, it has been shown that miRNAs can interact with targets via
sequences located in regions such as 5'UTR or CDS (Hausser et al.
\protect\hyperlink{ref-hausser_analysis_2013}{2013}; Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). These studies
also showed that binding location could indicate functionality of
miRNA:target interaction or be effective on abundance of targets(TOFIX
abundance etkisi derken, bazen azaltabilir, bazen arttırabilir,
manasında mı? answer: evet hocam anlaşılmıyorsa değiştirebilirim). It
has been shown that miRNAs exhibit repressive activity via, 6-8 nt long
sequence that is perfectly complementary with targets, seed region at
the 5' end of miRNAs (Bartel
\protect\hyperlink{ref-bartel_micrornas:_2009}{2009}; Grimson et al.
\protect\hyperlink{ref-grimson_microrna_2007}{2007}). On the other hand,
some researchers have reported that seed sequence of miRNA can have
mismatches or bulged/wobble nucleotides (Chi, Hannon, and Darnell
\protect\hyperlink{ref-chi2012alternative}{2012}) and may locate in
region other than 5'end of miRNAs (Hafner et al.
\protect\hyperlink{ref-hafner_transcriptome-wide_2010}{2010}; Helwak et
al. \protect\hyperlink{ref-helwak_mapping_2013}{2013}). On top of all
these factors, abundance of miRNAs and targets and miRNA:target ratio in
cells predominantly affect efficiency of miRNA:target interaction (Arvey
et al. \protect\hyperlink{ref-arvey_target_2010}{2010}; Bosson, Zamudio,
and Sharp \protect\hyperlink{ref-bosson_endogenous_2014}{2014}; Denzler
et al. \protect\hyperlink{ref-denzler_assessing_2014}{2014}).

As it is possible for miRNAs to suppress multiple targets, an individual
mRNA molecule can also be targeted by multiple miRNAs. In that case, the
targeted mRNAs exhibit competitor behavior, that is hypothesized as
competing endogenous RNAs (ceRNAs) (Ala et al.
\protect\hyperlink{ref-ala_integrated_2013}{2013}; Cesana and Daley
\protect\hyperlink{ref-cesana_deciphering_2013}{2013}), against their
miRNAs. Briefly, Ala et al.~have explained the ceRNA hypothesis as
disturbance of the other target when one of the targets on a
steady-state system that included one miRNA and two target was perturbed
with expression change (Ala et al.
\protect\hyperlink{ref-ala_integrated_2013}{2013}). Regarding
interaction between miRNAs and their target in a cell, explaining and
predicting results of an individual perturbation is difficult due to
complexity of interactions. Various computational and experimental
studies have tackled the problem of unraveling ceRNA:miRNA interactions.
It has been observed that when abundance of one of the targets of
miR-122 was increased, the other targets' expression also slightly
increased as a result of decreasing repressive activity of miR-122 on
them (Denzler et al.
\protect\hyperlink{ref-denzler_assessing_2014}{2014}). Bosson et
al.~have developed a mathematical model for changes on total target pool
concentration after grouping targets according to affinity and
demonstrated that miRNA activity correlated with affinity between miRNA
and target (Bosson, Zamudio, and Sharp
\protect\hyperlink{ref-bosson_endogenous_2014}{2014}). Cooperative
efficiency of miRNAs as well as competitor behaviors of targets were
also studied and it has been demonstrated to be crucial for regulating
available mRNA levels of targets (Denzler et al.
\protect\hyperlink{ref-denzler_impact_2016}{2016}). MiRNA:target
interactions have been modeled as stoichiometric and catalytic
mechanisms and Figliuzzi et al.~have recommended handling models in
network context (Figliuzzi, Marinari, and De Martino
\protect\hyperlink{ref-figliuzzi_micrornas_2013}{2013}). The model that
can explain miRNA target interaction through topological features has
been applied at bipartite network by Nitzan et al.~(Nitzan et al.
\protect\hyperlink{ref-nitzan_interactions_2014}{2014}). Robinson and
Henderson applied the model that handles miRNA:target direct and
indirect interactions via common targeting miRNA of genes and target of
miRNAs, at bipartite network. It has been demonstrated that all miRNAs
and targets in the network can interact with each other through common
miRNAs and genes, without interaction between the same type of nodes
(Robinson and Henderson
\protect\hyperlink{ref-robinson_modelling_2018}{2018}).Associated genes
that are targets of the same miRNAs have been found with help of
correlation of gene expression changes in recent algorithm (Markus List
\protect\hyperlink{ref-markus_list_sponge_2017}{2017}). List et al.~have
specified that their approach can be useful for ceRNA studies and
published their approach as an R package.

@@

Initially, the network is assumed in steady-state (\autoref{fig1}a)
condition and needs least one trigger for initiating calculations. The
trigger can be a change in expression level of one or more genes
(\autoref{fig1}b). After a trigger, the network undergoes iterative
cycle of calculations at each of which distribution of miRNA in local
neighborhood is recalculated (\autoref{fig1}c). Based on new miRNA
distribution, expression level of each node (i.e.~ceRNA) is updated.
This update results in change of expression value of common target (G4)
in system. In this case, the common target acts as trigger for the other
group (the targets interacted with M2 miRNA) in the network. Due to
common targeted elements, the change in one neighborhood spreads to
other neighborhoods (\autoref{fig1}d), consequently have potential to
effect whole network due to ``ripple effect''.

@@

We have applied our model in a real dataset for which experimental
measurements of various factors were available. Expression levels of
miRNA and genes in tumor and normal tissue of single patient are
retrieved from The Cancer Genome Atlas
\href{https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga}{TCGA}(\#TODO).
High-throughput experimental datasets which are provided miRNA:gene
target pairs with interaction factors (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). We have combined
miRNA and gene expression datasets via miRNA:target gene dataset
retrieved from \ldots{} (\#FIX name of database and its citation)
answer-\textgreater{}suplementary datasets of (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). Detailed
description of network construction and its code is available in
Supplementary data (\#TODO burada link nasıl olacak, word dosyası veya
PDF dosyası ismi mi vermemiz mi gerekiyor, yoksa sadece Sup Data
denilmesi yeterli mi? ilgili dosya: TCGA\_E9-A1N5\_article.Rmd). ABCC1
gene, over-expressed in tumor tissue, was selected as trigger for
simulation of integrated dataset. After simulation of network, we have
compared simulation results and tumor tissue expression levels.

@@

We have developed a network-based approach to assess effects of
expression level changes in competitive ceRNA regulation. The basic mode
of miRNA repression activity has been based on miRNA and target
abundance in various researches (Arvey et al.
\protect\hyperlink{ref-arvey_target_2010}{2010}; Denzler et al.
\protect\hyperlink{ref-denzler_assessing_2014}{2014}). Our approach can
effortlessly calculate effects of expression changes when abundance
levels of miRNAs and targets is only available factor. In sample network
given in Figure (\autoref{fig1}), after an increase in expression level
of a gene (G2), expression values of other genes also changed due to
redistribution of miRNA among its targets. Previous studies have shown
that if a gene abundance increases in ceRNA system, expression levels of
genes targeted by shared miRNA are also affected (Lai, Wolkenhauer, and
Vera \protect\hyperlink{ref-lai_understanding_2016}{2016}; Salmena et
al. \protect\hyperlink{ref-salmena_cerna_2011}{2011}; Tay, Rinn, and
Pandolfi \protect\hyperlink{ref-tay_multilayered_2014}{2014}). It was
observed that expression levels of primary neighborhoods which do not
interact with another miRNA of the trigger gene change in relation to
their expression. However, expression level change of common target (G4)
in network has been observed different from G1 that has the same
expression value with G4 target at initial conditions because more than
one miRNA repress this target (\autoref{fig1}d) (\#FIX last sentence
needs some clarification, after ``but cooperative\ldots{}'').

@@

In regulation of miRNA target sample system (\autoref{fig3}), miRNA (M1)
repressive efficiency on the primary triggering gene (G2) is low in
steady-state. So it has been observed that the regulatory activities of
miRNA on the targets are weak after the increase of miRNA target G2.
When model is triggered with two fold increase in expression level of
common target (G4), the changes of other gene expressions have observed
more prominent \href{}{}. Furthermore, change in expression level of
target gene that has strong miRNA repression efficiency resulted in
evident perturbation in network. On the other hand, it was observed that
Gene2 was weakly affected from change of Gene4 expression because of its
weak interaction factors. Since perturbation efficiency of each gene is
different, we have developed a function which screens each gene in the
network for their perturbation efficiencies. \textbf{(see details in
Supplementary File(fig1\_2app.Rmd)}. On the other hand, when we applied
the method on the minimal dataset, Gene4 has been found to be the most
efficient element in terms of number of perturbed elements and miRNA2
(M2) has been found to be causing the highest mean in expression
changes. {[}{]}{[}\#TODO2{]}

We used our model to simulate a real dataset which contains thousands of
genes and hundreds of miRNAs. Our model can successfully simulate
perturbations in such large network despite complex behaviors and
struggle to reach steady-state \textbf{(see details in Supplementary
File(TCGA\_E9-A1N5\_article.Rmd)}. Simulations show that change in
expression level of single gene has potential to effect whole network,
perturbing almost all nodes. These observations are in accordance with
competing endogenous RNA hypothesis where genes targeted with many
common miRNAs subsequently transmit perturbation to neighboring groups.

Network based approaches for analyzing miRNA:target interactions have
been developed in earlier studies. An initial attempt demonstrated the
ceRNA crosstalk in a network-like minimal interaction structure with
concentrations of ceRNA and miRNAs (Figliuzzi, Marinari, and De Martino
\protect\hyperlink{ref-figliuzzi_micrornas_2013}{2013}). Next, a network
based kinetic model integrating miRNA and target rates of transcription,
degradation, binding and unbinding was developed (Nitzan et al.
\protect\hyperlink{ref-nitzan_interactions_2014}{2014}) using high
throughput experimental dataset about miR-92a depletion (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}). It was demonstrated
that distant ceRNAs can interact with each other via indirect links, and
the interactions are effected depending on distance between ceRNAs or
topological features of network (Nitzan et al.
\protect\hyperlink{ref-nitzan_interactions_2014}{2014}). More recently,
an approach to detect ceRNA interaction by using the miRNA expression,
gene expression and common miRNAs between gene targets was developed
(Markus List \protect\hyperlink{ref-markus_list_sponge_2017}{2017})
which was effective in analyzing genes through miRNAs. Based on the
observations that a miRNA can exhibit strong functionality to a target
but may not against an other, the authors have concluded that existing
miRNA based approach may not be suitable for understanding regulations
of ceRNA interactions.

{[}\#Add{]}More recently, Silveira et al.~have developed miRmapper
package for using with R programme (Silveira et al.
\protect\hyperlink{ref-da2018mirmapper}{2018}). Researchers have
utilized the adjacency matrix to associate miRNAs using differentially
expressed genes and found similarities of miRNAs with this approach.
Authors who utilized the approach in dataset of bladder cancer cell line
have identified significant miRNAs in network, and miRNAs which work
synergistically.

{[}\#Add{]} Today, there are many of tools that provide predicted or
experimentally validated miRNA:target interaction dataset. The datasets
that contain weak, proteomic analysis like western blot, and strong,
crosslinking and immunoprecipitation CLiP, experimentally evidence are
more reliable sources than prediction datasets. While in experimentally
weak supported datasets it is not known that indirect or direct target
of miRNAs in cell, high-throughput methods ensure handling of AGO
(Argounate Protein) interacted miRNAs. For this reason we have preferred
to utilize high-throughput datasets in our application. Especially, we
collected datasets that depend on chimeric (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}) reading of miRNA:
target pairs from among these datasets because they contain exact
complementary seed regions.

{[}\#Add{]} Although these experimental sources include exact binding
sites, they do not provides which miRNA is functional or not (Liu and
Wang \protect\hyperlink{ref-liu2019prediction}{2019}). We have used the
features of bound miRNA:target pairs and scaled these parameters inside
groups that is consisted according to, ceRNA (Competing Endogenous RNA),
genes targeted by the same miRNA. In scaling process, we specified the
interaction factors as numeric values in accordance with previous
studies. Energy values in miRNA:target pairs are represented by
high-throughput studies (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}) which are utilized
in this study. On the other hand, we have specified the other
interaction factors, seed type and location of binding region on the
target, as numeric values based on the previous studies.(Grimson et al.
\protect\hyperlink{ref-grimson_microrna_2007}{2007}) have compared the
seed types' effect on target repression with few miRNA had canonical
seed pairing in their study. Additionally, (Bartel
\protect\hyperlink{ref-bartel_micrornas:_2009}{2009}) and (Betel et al.
\protect\hyperlink{ref-betel2010comprehensive}{2010}) have studied on
functional and non-functional seed interactions. Based on results of
these studies we have arranged seed types of miRNA:target interactions
as numeric values. We also have redefined location of binding region on
the target as numeric values, based on studies of (Hausser et al.
\protect\hyperlink{ref-hausser_analysis_2013}{2013}) and (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}). With this process,
we have handled this entegrated dataset in context of competitor
behaviours and functionality of interactions.

{[}\#Add{]}The main advantage of the method, ceRNAnetsim package, is
providing various functions which contain optionally arguments. So,
users can run the functions in accordance with their own data. For
example, the functions can work without interaction factors or with a
simple interaction factor that is specified by user. Furthermore, it can
also be operated according to a different parameter(s) provided by the
user.In this study, we have preferred to run model, shown in
\emph{Multifactorial calculations in miRNA:target network} section.

{[}\#TODO answer{]} We have observed that the results of simulation with
ABCC1 gene, \emph{Multidrug resistance-associated protein 1}, which is
one of the most significant factor to develop resistance against
chemotherapotic agents(Atalay, Demirkazik, and Gunduz
\protect\hyperlink{ref-atalay2008role}{2008}; Lu et al.
\protect\hyperlink{ref-lu2015microrna}{2015}; Atalay et al.
\protect\hyperlink{ref-atalay2006multidrug}{2006}) is not convenient
with tumor tissue expression values, when we compared the tumor gene
expression dataset retrieved from
\href{https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga}{TCGA}
and simulation results obtained via our approach. Simulation results
were not expected to be consistent with tumor tissue expression values
because large number of abnormal regulations were observed in these
tissues. So, it might not said that a single gene is responsible for
entire regulations in cancer cells. Besides, the all other factors such
as up/down regulation of miRNAs or other genes were ignored. Therefore,
it would be more useful to test our approach with an ideal dataset that
includes entire informations of gene expression values before and after
the regulation of a known gene at existing tissue, and miRNA expression
values of this tissue at initial conditions.

@@

\leavevmode\hypertarget{ref-atalay2006multidrug}{}%
Atalay, Can, Ismet Deliloglu Gurhan, Cigdem Irkkan, and Ufuk Gunduz.
2006. ``Multidrug Resistance in Locally Advanced Breast Cancer.''
\emph{Tumor Biology} 27 (6): 309--18.

\leavevmode\hypertarget{ref-atalay2008role}{}%
Atalay, C, A Demirkazik, and U Gunduz. 2008. ``Role of Abcb1 and Abcc1
Gene Induction on Survival in Locally Advanced Breast Cancer.''
\emph{Journal of Chemotherapy} 20 (6): 734--39.

@@

\leavevmode\hypertarget{ref-betel2010comprehensive}{}%
Betel, Doron, Anjali Koppal, Phaedra Agius, Chris Sander, and Christina
Leslie. 2010. ``Comprehensive Modeling of microRNA Targets Predicts
Functional Non-Conserved and Non-Canonical Sites.'' \emph{Genome
Biology} 11 (8): R90.

@@

\leavevmode\hypertarget{ref-chi2012alternative}{}%
Chi, Sung Wook, Gregory J Hannon, and Robert B Darnell. 2012. ``An
Alternative Mode of microRNA Target Recognition.'' \emph{Nature
Structural \& Molecular Biology} 19 (3): 321.

@@

\leavevmode\hypertarget{ref-liu2019prediction}{}%
Liu, Weijun, and Xiaowei Wang. 2019. ``Prediction of Functional microRNA
Targets by Integrative Modeling of microRNA Binding and Target
Expression Data.'' \emph{Genome Biology} 20 (1): 18.

\leavevmode\hypertarget{ref-lu2015microrna}{}%
Lu, Lin, Fang Ju, Hui Zhao, and Xuezhen Ma. 2015. ``MicroRNA-134
Modulates Resistance to Doxorubicin in Human Breast Cancer Cells by
Downregulating Abcc1.'' \emph{Biotechnology Letters} 37 (12): 2387--94.

@@

\leavevmode\hypertarget{ref-da2018mirmapper}{}%
Silveira, Willian da, Ludivine Renaud, Jonathan Simpson, William Glen,
Edward Hazard, Dongjun Chung, and Gary Hardiman. 2018. ``MiRmapper: A
Tool for Interpretation of miRNA--mRNA Interaction Networks.''
\emph{Genes} 9 (9): 458.
